\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-10-28 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex

\textit{where $d$ denotes the US district of entry and $k$ denotes an HS10 product} ($f_{ikd}$ being the transport costs). \textit{The
identification of $\beta_{ik}$, in this case, would rely on the across HS10 product and
district-of-entry variation in $f_{ik}$ and $p_{ik}$. Estimating the above equation would
obviously require that the authors do not aggregate up the raw Census data
across all districts and all 10-digit products pertaining to the same 5-digit category. [...] The first advantage of this so-called Hummel's approach is that the above regression can be estimated separately for various country-industry pairs, without imposing Assumptions (a) and (b) outlined above.}

\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummel's. We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy''). As we now show in this Subsection, the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs. The share of additive costs can hence be uncovered by regressing transport costs on unit fas prices.

Now coming to the technical side, we thank the referee for his/her relevant suggestion of an alternative estimation method. This drove us to question our own empirical specification deeper. In the end, we decided to maintain our empirical strategy in the revised version, because its advantages seem to overweight the possible costs, and the alternative that the referee suggests does not go without a few important limitations. We develop each line of argument in our answer. Yet, because the referee's estimation method is also an interesting alternative, we added a new section in the Online appendix devoted to it (Section D.3). After reading our answer, we hope to convince you that we made the right choice.

\paragraph{Estimating the share of additive costs: Highlighting some requirements} Our estimated equation relies on non-linear estimation methods, such as Non-Linear Least Squares. However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators. This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e. $\tau \geq 1$ and $t \geq 0$, or $0 \leq  \beta \leq 1$. Should we relax these restrictions, standard linear, least squares estimates often deliver negative, meaningless estimates. In this respect, implementing the referee's method (see below) does not dispose of the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces -- though the referee’s method is indeed less computational intensive). Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now presented in footnote \textbf{12 page 8}. \textbf{XXXXVérifier à la fin? GD}\smallskip

@@

\begin{itemize}
\item \textit{The first advantage of this so-called Hummel's Approach is that the above
regression can be estimated separately for various country-industry pairs, without
imposing Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.} \\

\textbf{Our answer:}

First, as already discussed in our letter, the separability assumption does not seem to be a strong assumption. This is based on the conclusion drawn from the robustness check on a reduced sample. If it is smaller in terms of observations, it yet remains quite large in terms of trade coverage, as the selected countries and sectors (at the 3-digit level) constitute 68\% of the total value of flows (on a yearly basis). As we conclude at the end of Section 3.3.1, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.

Second, the referee's method imposes a drastic selection of countries / sectors to be included in the estimation. Specifically, since the estimation method is run at the sector-origin country level (on top of being year-mode specific), the number of flows within each country-sector pair should be large enough to more than cover the number of $k$- product and $d$- entry port fixed effects. This substantially reduces the number of point estimates, as it can be inferred from Table \ref{tab_comp_referee1_baseline} reported later in our answer and in the Online Appendix, \textbf{Table D.1, Section D.3} ).

@@

\item  \textit{The third advantage is that, by adopting this approach, the comparison between the paper [...] and those in Hummels (2007) would become more transparent.}\\
     \textbf{Our answer:}
     We agree with the referee that the comparison with the literature (\citealp{hummels2007}) in particular) was not straightforward for the reader in the submitted version. We did our best to make the comparison clearer in the revised version, yet in a way consistent with our estimation method. In contrast to the submitted version, the paper now starts from the relation linking transport costs to the unit price specified in Hummels (2007):
    \begin{equation}
    f_{ikt} = X_{is(k)t}\widetilde{p}_{ikt}^{-\beta_{ikt}} \label{eq:Hummels}
    \end{equation}

     with $f_{ikt} = \frac{p_{ikt}}{\widetilde{p}_{ikt}} -1 $ the transport costs measure, $p_{ikt}$ ($\widetilde{p}_{ikt}$) the cif (fas) price and $\beta_{ikt}$ the price-elasticity of transport costs, with $i$ the origin country, $k$ the product and $t$ for time. As we now show in the paper (Section 2.2), this also corresponds to the share of additive costs in total transport costs:\footnote{For sake of expositional purpose, the reasoning is made without any $i,k,t$ specification in the revised paper. We directly report them here to ease the comparison with the referee's method afterwards.}

     \begin{equation}
     \beta_{ikt} = \frac{\frac{t_{is(k)}}{\widetilde{p}_{ikt}}}{\tau_{is(k)t}-1+\frac{t_{is(k)t}}{\widetilde{p}_{ikt}} } \label{eq:beta_TC}
     \end{equation}

     Equation (\ref{eq:Hummels}) lies at the root of Hummels's (2007) method. In contrast to our method though, \cite{hummels2007} estimates Equation (\ref{eq:Hummels}) on a panel basis and at the sectoral $s(k)$ basis, assuming a mode-specific $\beta$ invariant over time/sector/origin country. Rather than Equation (\ref{eq:Hummels}), this amounts starting from the functional form: $f_{ikt} = X_{is(k)t}\widetilde{p}_{ikt}^{-\beta}$.

     Two alternative strategies are possible. Let us start with our interpretation of the referee's method. It starts from the functional form linking transport costs and the fas price as specified in the Equation (\ref{eq:Hummels}). Taken in log (on a mode/yearly basis), it is written as:
     $$\ln f_{ikd} = \beta_{is(k)}\ln \widetilde{p}_{ikd} +\text{Controls}_{ikd}+ \varepsilon_{ikd}$$

@@

     \noindent with $d$ the district of entry. From this, one can then recover the levels of additive / multiplicative transports costs. Denoting $\widehat{\beta}_{is(k)}$ the estimated $\beta$ for a given sector-country $i,s(k)$, one can indeed solve the following two-equation system:

\begin{eqnarray}
p_{ik} &=& \tau_{is(k)}\widetilde{p}_{ik} +t_{is(k)} \label{eq:system1}\\
\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}+ t_{is(k)}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
\end{eqnarray}

\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode). With 2 equations and 2 endogenous variables, the system can be solved.

@@

 Alternatively, our method rather starts from the definition of transport costs as:
     $$\ln f_{ik} = \ln\left(\tau_{is(k)} -1 + \frac{t_{is(k)}}{\widetilde{p}_{ik}}\right) +\varepsilon_{ik}$$
\noindent from which we deduce the additive/multiplicative costs $\widehat{t}_{is(k)}$, $\widehat{\tau}_{is(k)}$ (on a mode/yearly basis). Then, relying on the Hummels' equation, one can deduce the share of additive costs in total transport costs $\beta_{ik}$ through Equation (\ref{eq:beta_TC}), again on a mode/yearly basis.

In this respect, the referee's method and ours are equivalent, in that they both allow to compute the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time and sectors (and transport mode). Put it differently, our method is consistent with Hummel's methodology - provided that it is adequately reported as such, as we do now.
In comparison with \cite{hummels2007}, one supplementary advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector. Section 4 ([Section 3] of the submitted version) explores the importance of allowing for such a variability in accounting for the sources of the transport costs time trend. We agree that this was not sufficiently transparent in the submitted version.
We hope that the revised version now meets the referee's expectations on this point.

@@

\end{itemize}

\paragraph{b) The costs are high}
\begin{itemize}
\item[Concern 1] The suggested estimation strategy implies having much less data to exploit, for two reasons.
\begin{enumerate}
\item Information about the port of entry are only available since 1989, which is already smaller than our full sample starting
in 1974. On top of that, because of the Covid situation, the US Bureau of Census was only able to send them the CDs relative to the years 1997-1999 and 2001-2019. Implementing the referee's method would hence necessarily reduce the time coverage of our analysis by more than 20 years (skipping the 1974-1996 years in particular). In our view, the historical coverage is interesting per se, as it provides useful insights about how transport costs have evolved over time. Eliminating this dimension of the paper would be detrimental to its value-added.
\item As underlined before, the method’s regressions are run country by country, and by 3-digit sector, exploiting the variability within each country-3d sector across 10-d sub-sectors and ports of entry. Yet, it appears that for many couples (country, 3-digit sector), there is too few variability across sub-sectors or ports en entry given the number of fixed effects included in the regression, such that estimation can not be run. This can be seen comparing the number of observations by year/ transport mode with our method / with the referee's method reported in Table \ref{tab_comp_referee1_baseline}. Put it differently, this methodology discards country-sector pairs that export a limited range of goods to the US and/or which arrive in the USA through the same ports of entry. In this respect, the induced selection bias may reduce the general scope of the transport costs estimates. This concern might yet be mitigated by noticing that the covered value of trade flows is only slightly lower. If it reduces the sample of sectors and origin countries covered, this method preserves the majority of trade flows in value.

    \begin{table}[htbp]
	\caption{Comparison 2005-2013}
	\begin{center}		
		\input{../../revised_article/Online_Appendix/Comp_baseline_referee1_reshaped.tex}
		\input{../../revised_article/Online_Appendix/Comp_baseline_referee1.tex}

@@

{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: Mean values reported for the 2005 and 2013 values for $\beta$. Time trend coefficient is the annual growth rate.}}}
\end{center}
	\label{tab_comp_referee1_baseline}%
\end{table}%
   \end{enumerate}

Following the referee's advice, we thus run this alternative estimation method as a robustness check. Table \ref{tab_comp_referee1_baseline} reports the estimation results, by transport mode, in the Column ``Alternative'', along with the results obtained with our baseline method over the same period 2005-2013. As noted above, the partners-sectors covered are lower. Regarding the share of additive costs $\beta$, the comparison between the two methods goes in opposite directions depending on the transport mode. While the estimated $\beta$ is lower for maritime transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport. In any case yet, the share of additive costs remains substantial, around 40\% on average over the period. In this respect, it confirms the main message of our paper shedding light on the importance of the additive component in international transport costs.

### 2021-10-28 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex

These three points indeed deserve careful consideration. We provide separate answers to each of them below.

%\subsubsection{On the justifications for our empirical approach}

\subsubsection{Concern (i): About the assumption of separability}
Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by \citet{Irrazabal_2015} to estimate the share of additive costs in a firm-level context. It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.

Further, we provide a robustness check for this separability assumption that $\tau_{ik} = \tau_i\tau_{k}$ and
 $t_{ik} = t_i +t_k$ in Section 3.3.1 of the paper (Table 2) \textbf{XXX CHECK AT THE END XXX}). We check the robustness of our results by re-running the estimation without the separability assumption. This comes at the cost of a substantial increase in the number of fixed in the regression. In 2019, for example, there are 14,016 country*sector pairs. Adding 28,032 fixed effects (number of country pairs $\times2$ for the two additive and multiplicative costs fixe effects) rather than ``only'' 836 ($(188+230)\times 2$). This is computationally not tractable.
% use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/data/hummels_tra.dta", clear
% keep if year==2019
% keep if mode=="ves"
% gen sector=substr(sitc2,1,3)
% gen blouk = iso_o+sector
% codebook blouk
 For this reason, we have decided to run the robustness check on a reduced sample.
 We select, for each year, the largest importers that form together at least 80\% of annual trade and the largest traded sectors that form together at least 80\% of annual trade.
 We keep in the sample all trade observations from these importers and in these sectors.
 This sample is smaller in terms of observations (2,125 for Air, 5,260 for Vessel on average over the period 1974-2019 (see Table 2), vs more than 30,000 on the complete sample, see Table 1), yet it remains quite large in terms of trade coverage (mean of 68\%).
% use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/results/comparaisons_various/stats_comp_baseline_non_séparé.dta", clear
% gen cover = cover_non_séparé/cover_baseline
% sum cover
\footnote{In the restricted sample at the 3-digit level, we have 25 sectors from 14 countries for Air transport, vs 212 sectors from 190 countries in the large sample; and 54 sectors from 20 countries for Vessel, vs 667 sectors from 192 countries in the complete sample). The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.}
 As we conclude at the end of Section 3.3.1, the trend pattern of the share of additive cost are very similar whether estimated under the separability assumption or not.
 %- see in particular Figures 6 and 7 on pp. 25-26 \textbf{XXX CHECK AT THE END XXX}. Given that the restricted sample covers 80\% of the total trade flows of the US economy, we hope to convince you that this result can be extended to the whole dataset.

@@

\subsubsection{Concern (ii): On the use of non-linear least squares (NLS)} As noted by the referee (and in the paper), it is true that relying on the non-linear least squares method is highly demanding computationally when the number of fixed effects is high. This constraint notably drove us to impose the separability assumption and retain $s=3$ as the relevant sectoral degree of aggregation.

The referee writes: \textit{A more natural approach is what the authors, at some point, refer to as the Hummels' Methodology. That is, one can alternatively estimate the share of the additive component as:}
$$\frac{t_{ik}}{ (\tau_{ik}-1)\tilde{p}_{ik} + t_{ik}} = \beta_{ik}$$

@@

\textit{where $\beta_{ik}$ is the elasticity of transport costs w.r.t. unit price. Given the authors' objective and the data they are using, $\beta$ can be separately estimated for each industry-country pair using the following regression:}

\begin{equation}
\ln f_{ikd} = \beta_{is(k)}\ln \tilde{p}_{ikd} + \text{Controls}_{ikd} +\epsilon_{ikd} \label{eq:estimation_ref1}
\end{equation}

\textit{where $d$ denotes the US district of entry and $k$ denotes an HS10 product} ($f_{ikd}$ being the transport costs). \textit{The
identification of $\beta_{ik}$, in this case, would rely on the across HS10 product and
district-of-entry variation in $f_{ik}$ and $p_{ik}$. Estimating the above equation would
obviously require that the authors do not aggregate up the raw Census data
across all districts and all 10-digit products pertaining to the same 5-digit category. [...] The first advantage of this so-called Hummels' approach is that the above regression can be estimated separately for various country-industry pairs, without imposing Assumptions (a) and (b) outlined above.}

\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummels'. We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy''). As we now show in this Subsection, the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs. The share of additive costs can hence be uncovered by regressing transport costs on unit fas prices.

@@

Now coming to the technical side, we thank the referee for his/her relevant suggestion of an alternative estimation method. This drove us to question our own empirical specification deeper. In the end, we decided to maintain our empirical strategy in the revised version, because its advantages seem to overweight the possible costs, and the alternative that the referee suggests does not go without a few important limitations. We develop each line of argument in our answer. Yet, because the referee's estimation method is also an interesting alternative, we added a new section in the Online appendix devoted to it (Section D.3). After reading our answer, we hope to convince you that we made the right choice.

\paragraph{Estimating the share of additive costs: Highlighting some requirements} Our estimated equation relies on non-linear estimation methods, such as Non-Linear Least Squares. However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators. This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e. $\tau \geq 1$ and $t \geq 0$, or $0 \leq  \beta \leq 1$. Should we relax these restrictions, standard linear, least squares estimates often deliver negative, meaningless estimates. In this respect, implementing the referee's method (see below) does not dispose of the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces -- though the referee’s method is indeed less computational intensive). Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote 12 page 8. \textbf{XXX Verifier à la fin? GD XXX }\smallskip

@@

\paragraph{Exploring Referee's alternative functional form \label{subsec:functional_form}}

The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007), yielding Equation (\ref{eq:estimation_ref1}) as estimation equation. By year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$). As previously noticed, running this estimation limits the time coverage as information is available at the HS-10 level and by ports of entry only since 2001. We have thus run referee's method over the years 2005-2013, at the $s=3, k=10$ sectoral/product levels, and reported the results in the Online Appendix.

Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$). We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the pros and cons of the referee's method versus ours, including those years for which the finest disaggregation level is $k=5$ (i.e., the original Hummels' dataset over 1974-1988, as the US Bureau of census import database starts in 1989 --yet being effectively available since 2001 only because of the Covid situation.) \textbf{Est-ec que cet argument peut vraiment être fait puisqu’on fait simplement 2005-2013 ?}

Despite its interest, implementing the referee's method uncovered some drawbacks, that in our view, outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps. First, we challenge the advantages that the referee's estimation method would bring in comparison with ours. Second, we emphasize the drawbacks induced by this method. \smallskip
%
%\paragraph{What we do}
%
%Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummels' original dataset from 1974-2004). We thus exploit the variability of transport costs between countries and between 3-digit sectors (conditional on a year and a transport mode, air or vessel). To sum up our methodology:
%\begin{itemize}
%\item Estimating Equation \ref{eq:equation0} provides us with estimates of $\hat{t}_i$, $\hat{t}_{s(k)}$, $\hat{\tau}_i$, $\hat{\tau}_{s(k)}$.
%\item We rebuild
%\begin{eqnarray*}
%\hat{t}_{is(k)} &= &\hat{t}_{i}+\hat{t}_{s(k)} \\
%\hat{\tau}_{is(k)} &= & \hat{\tau}_{i}\times\hat{\tau}_{s(k)}
%\end{eqnarray*}
%\item We deduce the weighted average value of each component $\bar{t}$, $\bar{\tau}$ by year and transport mode, the weighting scheme being the relative value of the flows for each $i,s$ flow, as well as the median, the maximum and the minimum values.
%\item With this in hand, we can (among other things) obtain the estimated share of additive costs in total transport costs for each $i,k$ flow:
%
%
%$$\hat{\beta}_{ik} = \frac{\frac{\hat{t}_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} $$
%\end{itemize}
%
%\paragraph{What the referee suggests to do}
%
%The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007). As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs. The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices. Specifically, the referee suggests to run the estimation based on Equation (\ref{eq:estimation_ref1}).
%
%Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10 digit level ($k$) and between ports of entry in the US ($d$).\footnote{The referee suggests to run his/her proposed estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$). We followed his suggestion but considering the 3-digit level. We made this choice in view of being able to compare the pros and cons of the referee's method versus ours. In Hummels' (2007) database, which we use until 2004, the finest degree of classification is at the 5-degree level until 2004. For comparability reasons over time, we adopt the same degree of classification $k=5$ for the following years (based on the original Census dataset). As a result, transport costs are estimated at the 3-digit level as benchmark estimate, exploiting heterogeneity at the 5-digit level (for a given origin country, and conditional on year and transport mode). This has driven us to retain $s=3$ when implementing the referee's method, in view of preserving comparability between both strategies.}

@@

\paragraph{a) The advantages of the referee's method are not as important as suggested}

\begin{itemize}
\item \textit{The first advantage of this so-called Hummels' Approach is that the above
regression can be estimated separately for various country-industry pairs, without
imposing Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.} \\

@@

\begin{eqnarray}
p_{ik} &=& \tau_{is(k)}\widetilde{p}_{ik} +t_{is(k)} \label{eq:system1}\\
\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}+ t_{is(k)}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
\end{eqnarray}

\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode). With 2 equations and 2 endogenous variables, the system might be solved.

 Alternatively, our method rather starts from the definition of transport costs as:
     $$\ln f_{ik} = \ln\left(\tau_{is(k)} -1 + \frac{t_{is(k)}}{\widetilde{p}_{ik}}\right) +\varepsilon_{ik}$$
\noindent from which we deduce the additive/multiplicative costs $\widehat{t}_{is(k)}$, $\widehat{\tau}_{is(k)}$ (on a mode/yearly basis). Then, relying on the Hummels' equation, one can deduce the share of additive costs in total transport costs $\beta_{ik}$ through Equation (\ref{eq:beta_TC}), again on a mode/yearly basis.

In this respect, the referee's method and our's are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time and sectors (and transport mode). Put it differently, our method is consistent with Hummels' methodology - provided that it is adequately reported as such. In comparison with \cite{hummels2007}, one supplementary advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector. Section 4 ([Section 3] of the submitted version) explores the importance of allowing for such a variability in accounting for the sources of the transport costs time trend. We agree that this was not sufficiently transparent in the submitted version. We hope that the revised version now fits the referee's expectations on this point.

@@

Following the referee's advice, we decided therefore to provide a full set of IV estimates to provide a clean assessment of the size of the potential bias. Section 4.3 in the revised version of the paper provides an overview of the results, while section B.4 in Appendix B provides a full presentation of the theoretical basis for the first-stage equation, as well as first-stage estimates. We follow earlier literature (see e.g. Caliendo and Parro, 2015, or Lashkaripour, 2017) by implementing a first-stage equation regressing $\widetilde{p}$ on custom duties coming based on tariffs at the product line, together with one-year lagged fas prices. First-stage estimates reported in section B.4 show that our main instrumental variable displays the right statistical properties, and that we can confidently re-inject the predictions arising from the first-stage equation for $\widetilde{p}_{ik}$ on the right-hand side of Equation 10 to produce a 2SLS-type of estimation. Figure 9 in the revised version of the paper reports our benchmark estimates by transport mode, together with their instrumented counterpart. In all cases, these estimates are very similar, when not identical. This supports that our benchmark estimates do not suffer any substantial biases arising from endogeneity concerns.

It should be noted that we performed this check on our main dataset (SITC 5 digit), and also at the HS 10 level, to handle simultaneously the referee's concern about aggregation issues. In the latter case, second-stage estimates are not reported for the sake of space, but do not change anything to previous conclusion - they are very, very similar to their non-IV, least squares counterparts. Needless to say that these results are available upon request from the referee.

\subsubsection{The aggregation problem}

\textit{\textbf{The aggregation problem}: The original annual Census data reports
trade at the origin country-HS10 product-district level of aggregation,
whereas the authors are aggregating up the data even further to the origin
country-HS5 industry-year level. Such an aggregation comes with strong implicit assumptions and sacrifices a lot of useful variation in the data.
The authors are motivating the aggregation by stating that the problem
would become computationally expensive without it. But this reasoning
brings us back to my original point that the authors can use the Hummels'
Methodology to circumvent the computational burden.}

It is true that the original annual Census data reports trade at the origin country-HS 10 product-district level of aggregation. Our initial choice of retaining $s=3$ at the sectoral level, $k=5$ at the product level was driven by the use of Hummels' dataset available over 1974-2004 with $k=5$ the finest degree of aggregation. We yet agree with the referee that jumping from $k=10$ to $k=5$ digits at the product level is detrimental to our ability to exploit useful variation in the data. After checking with the US Bureau of Census, data is only available since 1989, now available to 2019. As previously mentioned, due to the Covid situation, it was further only possible to receive the data over the years 1997-1999 and from 2001 onwards. In front of this arbitrage, we decided to maintain the case $s=3, k=5$ as baseline, but running estimation at the $s=3,k=10$ in the new Sub-Section 3.3.3 of Section 3.3 devoted to robustness analysis.

@@

Nevertheless, the point made by the referee is an important one and deserves careful investigation. We devoted a lot of thought to a potential switch from weight to quantity. Running the estimation considering quantities (i.e., units) rather than weight, still on the transport mode-yearly basis yet appears really uneasy in our context for several, non-trivial reasons:

\begin{enumerate}
\item In the original Hummels' database which we use from 1974 to 2004, we don't have information on quantity by transport mode. This is incompatible with our empirical strategy of estimation transport costs which is conditional on the transport mode, similarly as in \cite{hummels2007}. We could restrict observations to single-mode flows, but that would alter our sample in a non-trivial way, raising some potential selection biases. Investigating this possibility shows that this would reduce the total value of trade flows from around 32,000 billion dollars, to 7,000 billion dollars, i.e. a reduction of flows by a factor of 4.5.

@@

%XXX is chosen to match the share of firms exporting in Norway (38 percent, Moxnes, 2010), while

%IMO : The data cover all Norwegian non-oil exporters in 2004 and originate from customs declarations.

%\textbf{Comments about that}

%\paragraph{Piste 1} Consider his/her first suggestion. What is the idea? First, the link between gains from trade and trade-elasticity. Am I correct in saying that gains from trade are higher when trade is about low-elasticity goods? The idea being, if national goods are low substitutes, then here are larger gains from trading them. So, if composition effects (trade shifts between sectors) favor low-elasticity sectors, then one might expect high gains from trade. Hummels: trade composition effects matter as they partially offset the reduction in pure transport costs for both air and vessel. Over time, tendency to trade goods more costly to export everything else equal. So, if the referee's assertion is right (trade composition effects favor low-elasticity sectors), then increasing gains from trade. But, we disagree with Hummels' findings (his way of measuring things is inappropriate) and find that trade composition effects do not matter much, at least for air (for vessel, the composition effects matter more, but by amplifying the reduction in transport costs, ie towards goods that are less costly to export). Meaning that gains from trade are purely due to reduction in transport costs at the sector, but not from switch towards low-elasticity sectors. Identify one source of gains from trade, not two.

### 2021-10-28 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
Changement texte + online

\paragraph{Distribution of the $\beta$} As final check, we also investigate whether the varying pattern over time/country of origin/product is also robust to the separability assumption. With this aim, we report in Figure \ref{fig:histogram_beta_robustness}
the histogram of the distribution of the elasticity of transport costs to the export price ($\beta_{ikt}$, for air transport (top panel) and for maritime transport (bottom panel) under the baseline specification (Panels (A), (c))) and the non-separability scenario (Panels (B), (c)).

\begin{figure}[htbp]
\caption{Histogram of the share of additive costs in total costs (restricted sample)}
\label{fig:histogram_beta_robustness}
\begin{center}
\begin{tabular}{cc}
\includegraphics[width=14cm, height=10cm]{Etude_beta_pond.jpg}\\
\multicolumn{2}{l}{{\footnotesize Notes: Distribution of $\beta_{ikt}$ over the triplet (year-product-country), weighted by share of yearly value of flow.}}\\
\end{tabular}
\end{center}
\end{figure}

\textbf{Je ne comprends pas ce paragraphe. La footnote semble être ancienne. Mais en quoi est-ce que les histogrammes nous aident ? Éventuellement à dire qu’il y a moins de solutions de coin lorsqu’on fait l’estimation en séparé ? G.D.}For both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, country of origin, product) appears smoothly distributed over the interval $[0,1]$, whatever the assumption regarding the specification of fixed effects.
Relaxing the assumption that fixed effects are separable between the two country and product dimensions does not alter qualitatively our conclusion, that the elasticity of transport costs to the export price (or the share of additive costs in total transport costs), is strongly varying over time, countries of origin and products.
\footnote{As a consequence, this also provides an indirect robustness check to the results regarding the trend decomposition provided in Section 4 of the paper, which establish that the decline of ``pure'' transport costs is a main reason for the decrease of overall transport costs observed over time.}

@@

>>>>

@@

\textbf{CHANGER LEGENDE PAS TRES BEAU}
